\Large{Learning to Pivot with Adversarial Networks}

### 2016-10-31 tex/paper.tex
conclusions (work in progress)

Many inference problems involve data generation processes that are not uniquely
specified or are uncertain in some way. In a scientific context, the presence of
several plausible data generation processes is often associated to the presence
of systematic uncertainties. Robust inference is possible if it is based on a
pivot -- a quantity whose distribution is invariant to the unknown value of the
nuisance parameters that parametrize the family of data generation processes. In
this work, we introduce an adversarial training procedure to enforce the pivotal
property on a predictive model, which allows one to tune the tradeoff between
power and robustness. Under the assumption of existence, we derive theoretical
results proving that the proposed procedure converges towards a predictive model
that is both optimal and independent on the nuisance parameters, and demonstrate
the effectiveness of this approach with a toy and physics examples.

@@

\begin{figure*}
    \begin{center}
        \includegraphics[width=0.245\textwidth]{figures/f-plain.pdf}
        \includegraphics[width=0.245\textwidth]{figures/surface-plain.pdf}
        \includegraphics[width=0.245\textwidth]{figures/f-adversary.pdf}
        \includegraphics[width=0.245\textwidth]{figures/surface-adversary.pdf}
    \end{center}
    \vspace{-1cm}
    \caption{Toy example.
    (Left) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$
       when $f$ is built without adversarial training. The resulting densities
       are clearly dependent on the continuous parameter $Z$, indicating that $f$ is not pivotal.
    (Middle left) The corresponding decision surface, highlighting
       the fact that samples are easier to classify for values of $Z$ above to $\sigma$,
       hence partially explaining the dependency.
    (Middle right) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$ when $f$ is
       built with adversarial training, as outlined in Section~\ref{sec:method}.
       The resulting densities are now almost identical to each other, indicating only a
       small dependency on $Z$.
    (Right) The corresponding decision surface, illustrating how adversarial
       training bends the decision function vertically to erase the dependency on $Z$.
    }
    \label{fig:toy}
\end{figure*}

@@

>>>>

Many inference problems involve data generation processes that are not uniquely
specified or are uncertain in some way. In a scientific context, the presence of
several plausible data generation processes is often associated to the presence
of systematic uncertainties. Robust inference is possible if it is based on a
pivot -- a quantity whose distribution is invariant to the unknown value of the
(categorical or continuous)
nuisance parameters that parametrize the family of data generation processes. In
this work, we introduce a flexible adversarial training procedure to enforce the pivotal
property on a predictive model, which allows one to tune the tradeoff between
power and robustness. Under the assumption of existence, we derive theoretical
results proving that the proposed procedure converges towards a predictive model
that is both optimal and independent on the nuisance parameters, and demonstrate
the effectiveness of this approach with a toy and physics examples.

@@

\begin{figure*}
    \begin{center}
        \includegraphics[width=0.245\textwidth]{figures/f-plain.pdf}
        \includegraphics[width=0.245\textwidth]{figures/surface-plain.pdf}
        \includegraphics[width=0.245\textwidth]{figures/f-adversary.pdf}
        \includegraphics[width=0.245\textwidth]{figures/surface-adversary.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{Toy example.
    (Left) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$
       when $f$ is built without adversarial training. The resulting densities
       are clearly dependent on the continuous parameter $Z$, indicating that $f$ is not pivotal.
    (Middle left) The corresponding decision surface, highlighting
       the fact that samples are easier to classify for values of $Z$ above to $\sigma$,
       hence partially explaining the dependency.
    (Middle right) Conditional probability densities of the decision scores at $Z=-\sigma, 0, \sigma$ when $f$ is
       built with adversarial training, as outlined in Section~\ref{sec:method}.
       The resulting densities are now almost identical to each other, indicating only a
       small dependency on $Z$.
    (Right) The corresponding decision surface, illustrating how adversarial
       training bends the decision function vertically to erase the dependency on $Z$.
    }
    \label{fig:toy}
\end{figure*}

@@

In this work, we proposed a flexible learning procedure for building a
predictive model that is independent on continuous or categorical nuisance
parameters by jointly training two neural networks in an adversarial fashion.
From a theoretical perspective, we showed that the approach leads to a
predictive model that is both optimal and pivotal (if that models exists) or for
which one can tune the trade-off between power and robustness. From an empirical
point of view, we confirmed the effectiveness of our method on a toy and physics
example.

\glnote{Add future works.}
\glnote{Mention limitations of the current algorithm.}
